---
title: "AIM 2 Stats"
output: html_document
date: "2022-09-07"
---

## Test for Normality with Shapiro Test
#### TTIM data
```{r }
library(xlsx)
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "AL_TTIM_M")
shapiro.test(data$TTIM)
```

```{r }
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "AL_TTIM_A")
shapiro.test(data$TTIM)
```

```{r }
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "Day1TTIM_AL")
shapiro.test(data$TTIM)
```

```{r }
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "Day2TTIM_AL")
shapiro.test(data$TTIM)
```

```{r }
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "CR_TTIM_M")
shapiro.test(data$TTIM)
```

```{r }
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "CR_TTIM_A")
shapiro.test(data$TTIM)
```

```{r }
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "Day1TTIM_CR")
shapiro.test(data$TTIM)
```

```{r }
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "Day2TTIM_CR")
shapiro.test(data$TTIM)
```

```{r }
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "Day1TTIM")
shapiro.test(data$TTIM)
```

```{r }
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "Day2TTIM")
shapiro.test(data$TTIM)
```
#### %SAV data
```{r, %SAV}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "P_SAV")

shapiro.test(data$P_SAV)

data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "AL_P_SAV")
shapiro.test(data$P_SAV)

data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "CR_P_SAV")
shapiro.test(data$P_SAV)

```

#### TET data
```{r }
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "TET_AL_M")
shapiro.test(data$TET)
```

```{r }
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "TET_AL_A")
shapiro.test(data$TET)
```

```{r }
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "TET_Day1_AL")
shapiro.test(data$TET)
```

```{r }
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "TET_Day2_AL")
shapiro.test(data$TET)
```

```{r }
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "TET_CR_M")
shapiro.test(data$TET)
```

```{r }
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "TET_CR_A")
shapiro.test(data$TET)
```

```{r }
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "TET_Day1_CR")
shapiro.test(data$TET)
```

```{r }
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "TET_Day2_CR")
shapiro.test(data$TET)
```


## Stats Test
### AL %SAV
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "AL_P_SAV")

t.test(data$P_SAV ~ data$Age)
```

### CR %SAV
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "CR_P_SAV")

t.test(data$P_SAV ~ data$Age)
```


## Stats Test
### AL TTIM M
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "AL_TTIM_M")

library(dplyr)
group_by(data,Day) %>%
  summarise(
    count = n(),
    median = median(TTIM, na.rm = TRUE),
    IQR = IQR(TTIM, na.rm = TRUE))
 
# loading package for boxplot
library(ggpubr)
ggboxplot(data, x = "Day", y = "TTIM",
          color = "Day", palette = c("#FFA500", "#FF0000"),
          ylab = "TTIM", xlab = "Day")
 

# res <- t.test(data$TTIM[data$Day==1], data$TTIM[data$Day==2], paired = T)
# 
# res

## NEW STATS
res <- wilcox.test(TTIM ~ Day,
                   data = data,
                   paired = TRUE)
res
```
### AL TTIM A
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "AL_TTIM_A")


group_by(data,Day) %>%
  summarise(
    count = n(),
    median = median(TTIM, na.rm = TRUE),
    IQR = IQR(TTIM, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "Day", y = "TTIM",
          color = "Day", palette = c("#FFA500", "#FF0000"),
          ylab = "TTIM", xlab = "Day")

# res <- wilcox.test(TTIM ~ Day,
#                    data = data,
#                    exact = FALSE)
# res

## NEW STATS
res <- wilcox.test(TTIM ~ Day,
                   data = data,
                   paired = TRUE)
res
```
### AL TTIM Day 1
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "Day1TTIM_AL")


group_by(data,Age) %>%
  summarise(
    count = n(),
    median = median(TTIM, na.rm = TRUE),
    IQR = IQR(TTIM, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "Age", y = "TTIM",
          color = "Age", palette = c("#FFA500", "#FF0000"),
          ylab = "TTIM", xlab = "Age")

res <- wilcox.test(TTIM ~ Age,
                   data = data,
                   exact = FALSE)
res
```
### AL TTIM Day 2
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "Day2TTIM_AL")


group_by(data,Age) %>%
  summarise(
    count = n(),
    median = median(TTIM, na.rm = TRUE),
    IQR = IQR(TTIM, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "Age", y = "TTIM",
          color = "Age", palette = c("#FFA500", "#FF0000"),
          ylab = "TTIM", xlab = "Age")

res <- wilcox.test(TTIM ~ Age,
                   data = data,
                   exact = FALSE)
res
```
### CR TTIM M
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "CR_TTIM_M")


group_by(data,Day) %>%
  summarise(
    count = n(),
    median = median(TTIM, na.rm = TRUE),
    IQR = IQR(TTIM, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "Day", y = "TTIM",
          color = "Day", palette = c("#FFA500", "#FF0000"),
          ylab = "TTIM", xlab = "Day")

# res <- t.test(data$TTIM[data$Day==1], data$TTIM[data$Day==2], paired = T)
# 
# res

## NEW STATS
res <- wilcox.test(TTIM ~ Day,
                   data = data,
                   paired = TRUE)
res
```
### CR TTIM A
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "CR_TTIM_A")


group_by(data,Day) %>%
  summarise(
    count = n(),
    median = median(TTIM, na.rm = TRUE),
    IQR = IQR(TTIM, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "Day", y = "TTIM",
          color = "Day", palette = c("#FFA500", "#FF0000"),
          ylab = "TTIM", xlab = "Day")

# res <- t.test(data$TTIM[data$Day==1], data$TTIM[data$Day==2], paired = T)
# 
# res

## NEW STATS
res <- wilcox.test(TTIM ~ Day,
                   data = data,
                   paired = TRUE)
res
```

### CR TTIM Day 1
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "Day1TTIM_CR")


group_by(data,Age) %>%
  summarise(
    count = n(),
    median = median(TTIM, na.rm = TRUE),
    IQR = IQR(TTIM, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "Age", y = "TTIM",
          color = "Age", palette = c("#FFA500", "#FF0000"),
          ylab = "TTIM", xlab = "Age")

# res <- t.test(data$TTIM[data$Age=="M"], data$TTIM[data$Age=="A"], paired = T)
# 
# res

## New Test
res <- wilcox.test(TTIM ~ Age,
                   data = data,
                   exact = FALSE)
res
```
### CR TTIM Day 2
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "Day2TTIM_CR")


group_by(data,Age) %>%
  summarise(
    count = n(),
    median = median(TTIM, na.rm = TRUE),
    IQR = IQR(TTIM, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "Age", y = "TTIM",
          color = "Age", palette = c("#FFA500", "#FF0000"),
          ylab = "TTIM", xlab = "Age")

# res <- t.test(data$TTIM[data$Age=="M"], data$TTIM[data$Age=="A"], paired = T)
# 
# res

## New Test
res <- wilcox.test(TTIM ~ Age,
                   data = data,
                   exact = FALSE)
res
```

# Stats Test
### AL Bites M
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "AL_Bites_M")

group_by(data,Day) %>%
  summarise(
    count = n(),
    median = median(Bites, na.rm = TRUE),
    IQR = IQR(Bites, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "Day", y = "Bites",
          color = "Day", palette = c("#FFA500", "#FF0000"),
          ylab = "Bites", xlab = "Day")
 

# res <- t.test(data$TTIM[data$Day==1], data$TTIM[data$Day==2], paired = T)
# 
# res

## NEW STATS
res <- wilcox.test(Bites ~ Day,
                   data = data,
                   paired = TRUE)
res

#chisq.test(data$Bites, data$Day, correct = F) <- Not the correct test!!
```

### AL Bites A
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "AL_Bites_A")

group_by(data,Day) %>%
  summarise(
    count = n(),
    median = median(Bites, na.rm = TRUE),
    IQR = IQR(Bites, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "Day", y = "Bites",
          color = "Day", palette = c("#FFA500", "#FF0000"),
          ylab = "Bites", xlab = "Day")
 

# res <- t.test(data$TTIM[data$Day==1], data$TTIM[data$Day==2], paired = T)
# 
# res

## NEW STATS
res <- wilcox.test(Bites ~ Day,
                   data = data,
                   paired = TRUE)
res
```

### AL Bites Day 1
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "Day1Bites_AL")

group_by(data,Age) %>%
  summarise(
    count = n(),
    median = median(Bites, na.rm = TRUE),
    IQR = IQR(Bites, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "Age", y = "Bites",
          color = "Age", palette = c("#FFA500", "#FF0000"),
          ylab = "Bites", xlab = "Age")
 

# res <- t.test(data$TTIM[data$Day==1], data$TTIM[data$Day==2], paired = T)
# 
# res

## NEW STATS
res <- wilcox.test(Bites ~ Age,
                   data = data,
                   exact = FALSE)
res
```

### AL Bites Day 2
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "Day2Bites_AL")

group_by(data,Age) %>%
  summarise(
    count = n(),
    median = median(Bites, na.rm = TRUE),
    IQR = IQR(Bites, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "Age", y = "Bites",
          color = "Age", palette = c("#FFA500", "#FF0000"),
          ylab = "Bites", xlab = "Age")
 

# res <- t.test(data$TTIM[data$Day==1], data$TTIM[data$Day==2], paired = T)
# 
# res

## NEW STATS
res <- wilcox.test(Bites ~ Age,
                   data = data,
                   exact = FALSE)
res
```

### CR Bites M
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "CR_Bites_M")

group_by(data,Day) %>%
  summarise(
    count = n(),
    median = median(Bites, na.rm = TRUE),
    IQR = IQR(Bites, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "Day", y = "Bites",
          color = "Day", palette = c("#FFA500", "#FF0000"),
          ylab = "Bites", xlab = "Day")
 

# res <- t.test(data$TTIM[data$Day==1], data$TTIM[data$Day==2], paired = T)
# 
# res

## NEW STATS
res <- wilcox.test(Bites ~ Day,
                   data = data,
                   paired = TRUE)
res
```

### CR Bites A
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "CR_Bites_A")

group_by(data,Day) %>%
  summarise(
    count = n(),
    median = median(Bites, na.rm = TRUE),
    IQR = IQR(Bites, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "Day", y = "Bites",
          color = "Day", palette = c("#FFA500", "#FF0000"),
          ylab = "Bites", xlab = "Day")
 

# res <- t.test(data$TTIM[data$Day==1], data$TTIM[data$Day==2], paired = T)
# 
# res

## NEW STATS
res <- wilcox.test(Bites ~ Day,
                   data = data,
                   paired = TRUE)
res
```

### CR Bites Day 1
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "Day1Bites_CR")

group_by(data,Age) %>%
  summarise(
    count = n(),
    median = median(Bites, na.rm = TRUE),
    IQR = IQR(Bites, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "Age", y = "Bites",
          color = "Age", palette = c("#FFA500", "#FF0000"),
          ylab = "Bites", xlab = "Age")
 

# res <- t.test(data$TTIM[data$Day==1], data$TTIM[data$Day==2], paired = T)
# 
# res

## NEW STATS
res <- wilcox.test(Bites ~ Age,
                   data = data,
                   exact = FALSE)
res
```

### CR Bites Day 2
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "Day2Bites_CR")

group_by(data,Age) %>%
  summarise(
    count = n(),
    median = median(Bites, na.rm = TRUE),
    IQR = IQR(Bites, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "Age", y = "Bites",
          color = "Age", palette = c("#FFA500", "#FF0000"),
          ylab = "Bites", xlab = "Age")
 

# res <- t.test(data$TTIM[data$Day==1], data$TTIM[data$Day==2], paired = T)
# 
# res

## NEW STATS
res <- wilcox.test(Bites ~ Age,
                   data = data,
                   exact = FALSE)
res
```

```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "Day1TTIM")


group_by(data,group) %>%
  summarise(
    count = n(),
    median = median(TTIM, na.rm = TRUE),
    IQR = IQR(TTIM, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "group", y = "TTIM",
          color = "group", palette = c("#FFA500", "#FF0000", "#FF9999", "#FF9900"),
          ylab = "TTIM", xlab = "group")

kruskal.test(TTIM ~ group,
             data = data)

## There are significant differences between Months. Next is a post hoc test

PH <- dunnTest(TTIM ~ group,
               data=data,
               method = "bh")

PH
```

```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "Day2TTIM")


group_by(data,group) %>%
  summarise(
    count = n(),
    median = median(TTIM, na.rm = TRUE),
    IQR = IQR(TTIM, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "group", y = "TTIM",
          color = "group", palette = c("#FFA500", "#FF0000", "#FF9999", "#FF9900"),
          ylab = "TTIM", xlab = "group")

one.way <- aov(TTIM ~ group, 
    data = data)
summary(one.way)

TukeyHSD(one.way)
```

```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "P_SAV")


group_by(data,group) %>%
  summarise(
    count = n(),
    median = median(P_SAV, na.rm = TRUE),
    IQR = IQR(P_SAV, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "group", y = "P_SAV",
          color = "group", palette = c("#FFA500", "#FF0000", "#FF9999", "#FF9900"),
          ylab = "P_SAV", xlab = "group")

one.way <- aov(P_SAV ~ group, 
    data = data)
summary(one.way)

TukeyHSD(one.way)



kruskal.test(P_SAV ~ group,
             data = data)

## There are significant differences between Months. Next is a post hoc test

PH <- dunnTest(P_SAV ~ group,
               data=data,
               method = "bh")

PH
```

## Stats Test
### AL TET M
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "TET_AL_M")


group_by(data,Day) %>%
  summarise(
    count = n(),
    median = median(TET, na.rm = TRUE),
    IQR = IQR(TET, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "Day", y = "TET",
          color = "Day", palette = c("#FFA500", "#FF0000"),
          ylab = "TET", xlab = "Day")
 

# res <- t.test(data$TTIM[data$Day==1], data$TTIM[data$Day==2], paired = T)
# 
# res

## NEW STATS
res <- wilcox.test(TET ~ Day,
                   data = data,
                   paired = TRUE)
res
```

### AL TET A
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "TET_AL_A")


group_by(data,Day) %>%
  summarise(
    count = n(),
    median = median(TET, na.rm = TRUE),
    IQR = IQR(TET, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "Day", y = "TET",
          color = "Day", palette = c("#FFA500", "#FF0000"),
          ylab = "TET", xlab = "Day")
 

# res <- t.test(data$TTIM[data$Day==1], data$TTIM[data$Day==2], paired = T)
# 
# res

## NEW STATS
res <- wilcox.test(TET ~ Day,
                   data = data,
                   paired = TRUE)
res
```


### AL TET Day 1
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "TET_Day1_AL")


group_by(data,Age) %>%
  summarise(
    count = n(),
    median = median(TET, na.rm = TRUE),
    IQR = IQR(TET, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "Age", y = "TET",
          color = "Age", palette = c("#FFA500", "#FF0000"),
          ylab = "TET", xlab = "Age")

res <- wilcox.test(TET ~ Age,
                   data = data,
                   exact = FALSE)
res
```

### AL TET Day 2
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "TET_Day2_AL")


group_by(data,Age) %>%
  summarise(
    count = n(),
    median = median(TET, na.rm = TRUE),
    IQR = IQR(TET, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "Age", y = "TET",
          color = "Age", palette = c("#FFA500", "#FF0000"),
          ylab = "TET", xlab = "Age")

res <- wilcox.test(TET ~ Age,
                   data = data,
                   exact = FALSE)
res
```


### CR TET M
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "TET_CR_M")


group_by(data,Day) %>%
  summarise(
    count = n(),
    median = median(TET, na.rm = TRUE),
    IQR = IQR(TET, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "Day", y = "TET",
          color = "Day", palette = c("#FFA500", "#FF0000"),
          ylab = "TET", xlab = "Day")


res <- t.test(data$TET[data$Day==1], data$TET[data$Day==2], paired = T, alternative = "two.sided")

res
```

### CR TET A
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "TET_CR_A")


group_by(data,Day) %>%
  summarise(
    count = n(),
    median = median(TET, na.rm = TRUE),
    IQR = IQR(TET, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "Day", y = "TET",
          color = "Day", palette = c("#FFA500", "#FF0000"),
          ylab = "TET", xlab = "Day")


res <- t.test(data$TET[data$Day==1], data$TET[data$Day==2], paired = T, alternative = "two.sided")

res
```


### CR TET Day 1
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "TET_Day1_CR")


group_by(data,Age) %>%
  summarise(
    count = n(),
    median = median(TET, na.rm = TRUE),
    IQR = IQR(TET, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "Age", y = "TET",
          color = "Age", palette = c("#FFA500", "#FF0000"),
          ylab = "TET", xlab = "Age")


res <- t.test(data$TET[data$Age=="M"], data$TET[data$Age=="A"], paired = F, alternative = "two.sided")

res
```


### CR TET Day 2
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "TET_Day2_CR")


group_by(data,Age) %>%
  summarise(
    count = n(),
    median = median(TET, na.rm = TRUE),
    IQR = IQR(TET, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "Age", y = "TET",
          color = "Age", palette = c("#FFA500", "#FF0000"),
          ylab = "TET", xlab = "Age")


res <- t.test(data$TET[data$Age=="M"], data$TET[data$Age=="A"], paired = F, alternative = "two.sided")

res
```


# TWR habituation
```{r}
data <- read.xlsx("/Users/school/Documents/Lab Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/TWRhabituation.xlsx", sheetName = "TWR_Y_CR")
data$subject <- as.factor(data$subject)
data$Time <- as.factor(data$Time)

# ggboxplot(data, x = "Time", y = "Normalized.resp.", add = "point")
# 
# model = aov(Normalized.resp. ~ Time, data = data)
# summary(model)
# 
# 
# TukeyHSD(model) %>% view()

## Not sure how to use this data. after the above test and looking at the padj values in the Tukey test the only group that significantly habituated was the younger AL animals. Maybe statistics should not be used although I belive I did them wrong. Will try to come back to this later.

#http://agroninfotech.blogspot.com/2020/06/one-way-repeated-measures-anova-in-r.html

model = aov(formula = Normalized.resp. ~ Time + Error(subject/Time), data = data)
summary(model)

summary = data %>%
          group_by(Time) %>%
          summarize(Mean_Normalized.resp. = mean(Normalized.resp.),
                    SD_resp = sd(Normalized.resp.),
                    SE_resp = sd(Normalized.resp.)/sqrt(length(Normalized.resp.)),
                    n = n())
print(summary)

plot(summary$Mean_Normalized.resp.,
     type = 'o',
     xlab = 'Time',
     ylab = 'response')

pairwise.t.test(
          x = data$Normalized.resp.,
          g = data$Time,
          p.adjust.method = 'bonferroni'
          
) %>% view()
## The above code works a lot better but it is still not showing significance in any group except the AL younger samples
```

# CR vs AL Day 1
### TTIM in Younger
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "TTIM_Day1_CRvsAL_Y")


group_by(data,Diet) %>%
  summarise(
    count = n(),
    median = median(TTIM, na.rm = TRUE),
    IQR = IQR(TTIM, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "Diet", y = "TTIM",
          color = "Diet", palette = c("#FFA500", "#FF0000"),
          ylab = "TTIM", xlab = "Diet")

res <- wilcox.test(TTIM ~ Diet,
                   data = data,
                   exact = FALSE)
res
```

### TTIM in Aged
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "TTIM_Day1_CRvsAL_A")


group_by(data,Diet) %>%
  summarise(
    count = n(),
    median = median(TTIM, na.rm = TRUE),
    IQR = IQR(TTIM, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "Diet", y = "TTIM",
          color = "Diet", palette = c("#FFA500", "#FF0000"),
          ylab = "TTIM", xlab = "Diet")

res <- wilcox.test(TTIM ~ Diet,
                   data = data,
                   exact = FALSE)
res
```

### Bite Frequency in Younger
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "Bites_Day1_CRvsAL_Y")

group_by(data,Diet) %>%
  summarise(
    count = n(),
    median = median(Bites, na.rm = TRUE),
    IQR = IQR(Bites, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "Diet", y = "Bites",
          color = "Diet", palette = c("#FFA500", "#FF0000"),
          ylab = "Bites", xlab = "Diet")
 

res <- wilcox.test(Bites ~ Diet,
                   data = data,
                   exact = FALSE)
res
```


### Bite Frequency in Aged
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "Bites_Day1_CRvsAL_A")

group_by(data,Diet) %>%
  summarise(
    count = n(),
    median = median(Bites, na.rm = TRUE),
    IQR = IQR(Bites, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "Diet", y = "Bites",
          color = "Diet", palette = c("#FFA500", "#FF0000"),
          ylab = "Bites", xlab = "Diet")
 

res <- wilcox.test(Bites ~ Diet,
                   data = data,
                   exact = FALSE)
res
```

### TET in Younger
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "TET_Day1_CRvsAL_Y")


group_by(data,Diet) %>%
  summarise(
    count = n(),
    median = median(TET, na.rm = TRUE),
    IQR = IQR(TET, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "Diet", y = "TET",
          color = "Diet", palette = c("#FFA500", "#FF0000"),
          ylab = "TET", xlab = "Diet")

res <- wilcox.test(TET ~ Diet,
                   data = data,
                   exact = FALSE)
res
```

### TET in Aged
```{r}
data <- read.xlsx("/Users/school/Documents/Documents/UM_Lab_Work/AIM2_Caloric_Restriction/Paper/Results_and_Figures/AIM2data.xlsx", sheetName = "TET_Day1_CRvsAL_A")


group_by(data,Diet) %>%
  summarise(
    count = n(),
    median = median(TET, na.rm = TRUE),
    IQR = IQR(TET, na.rm = TRUE))
 
# loading package for boxplot
ggboxplot(data, x = "Diet", y = "TET",
          color = "Diet", palette = c("#FFA500", "#FF0000"),
          ylab = "TET", xlab = "Diet")

res <- wilcox.test(TET ~ Diet,
                   data = data,
                   exact = FALSE)
res
```



